---
title: "CITE_seq"
author: "douj"
date: "February 9, 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is used to show bindSC performance on aligning RNA and cytof data using CITE-seq data. 



```{r load library/packages}
source("../../bindsc/BiCCA.R")
source("../../bindsc/generics.R")

```

## Load PBMC dataset 

```{r load data, echo=FALSE}
bm <- readRDS(file="../../data/cite_seq.RDS")
summary(bm)
dim(bm$RNA)
dim(bm$Protein)
dim(bm$Z)

# downsampling 
N <- ncol(bm$RNA)
Y <- bm$RNA[,seq(1,N,10)]
X <- bm$Protein[,seq(1,N,10)]
Z0 <- bm$Z[,seq(1,N,10)]
```
```{r run bindSC }

source("../../bindsc/BiCCA.R")

X.clst <- as.character(bm$meta$celltype2)[seq(1,N,10)]
Y.clst <- as.character(bm$meta$celltype2)[seq(1,N,10)]


paraSel <- BiCCA_para_opt( X = X ,
             Y = Y, 
             Z0 =Z0, 
             tp="out",
             X.clst = X.clst,
             Y.clst = Y.clst,
             alpha.lst = seq(0,1,0.1), 
             K.lst = c(10),
             lambda.lst = seq(0,1,0.1),
             num.iteration = 50,
             tolerance = 0.01,
             save = TRUE,
             block.size = 1000
)


```



```{r, message = FALSE, warning=FALSE, results='hide',fig.height=3, fig.width=12}
source("../../bindsc/generics.R")
p1 <- paraSel_plot(paraSel)
p1

tiff("paraSel.tiff", width=12, height =3, res =300, units = "in")
print(p1)
dev.off()

```
# The decompostion has the best alignment/silhouette scores when alpha<0.5. The decompostion is relatively robust to 
# lambda exccept for alpha=0. Overall, we set alpha = 0.1 and lambda = 0.7 as the optimal parameter combination. 
# Next we will show results with alpha = 0.1 and lambda = 0.7

```{r, message = FALSE, warning=FALSE, results='hide',fig.height=5, fig.width=9}
source("../../bindsc/BiCCA.R")
source("../../bindsc/generics.R")
  res <- BiCCA(X = X,
               Y = Y, 
               Z0 =Z0, 
               alpha =0.1, 
               lambda = 0.7, 
               X.clst = X.clst,
               Y.clst = Y.clst,
               temp.path  = "./out",
               K = 10, 
               num.iteration = 50,
               tolerance = 0.01,
               save = TRUE)

p2 <- plotIteration(res)
p2

```
# The decompostion stops after the 13rd iteration. We will see how UMAP looks like when the iteration process stops

```{r, message = FALSE, warning=FALSE, results='hide',fig.height=6, fig.width=12}
alpha <- 0.1
lambda <- 0.7 
celltype <- as.character(bm$meta$celltype2[seq(1,N,10)])
umap_plt <- umap(rbind(res$u, res$r))
umap_plt  <- data.frame("UMAP1"=umap_plt$layout[,1],
                        "UMAP2"=umap_plt$layout[,2],
                        "celltype" = c(celltype, celltype),
                        "data" = c(rep("protein",length(celltype)),
                                   rep("RNA",length(celltype))))
  
xlim <- c(min(umap_plt$UMAP1), max(umap_plt$UMAP1))
ylim <- c(min(umap_plt$UMAP2), max(umap_plt$UMAP2))

p11 <- UMAP_plot(meta = umap_plt[umap_plt$data=="protein",], 
                color = "celltype", xlim = xlim, ylim = ylim)
p12 <- UMAP_plot(meta = umap_plt[umap_plt$data=="RNA",], 
              color = "celltype", xlim = xlim, ylim = ylim)

p11 <- p11 + ggtitle("Protein")
p12 <- p12 + ggtitle("RNA")

p3 <- ggarrange(p11,p12, ncol=2)
print(p3)

```
# Given the cell correspondence between RNA and Protein is known, we can validate whether the iteration process 
# has improved the gene score matrix Z. 
```{r, message = FALSE, warning=FALSE, results='hide',fig.height=6, fig.width=12}

p5 <- plot_geneScoreChange(X=X, Z0 = Z0, Z_impu = res$Z_est)

p5


```


# Show imputation accuracy after bindSC for each protein markers 
```{r, message = FALSE, warning=FALSE, results='hide',fig.height=6, fig.width=6}
cor1 <- diag(cor(as.matrix(t(X)),as.matrix(t(Z0)), method = "pearson"))
cor2 <- diag(cor(as.matrix(t(X)),as.matrix(t(res$Z_est)), method = "pearson"))
plt_dt <- data.frame("protein"=names(cor1),
                     "cor1"=cor1, "cor2"=cor2)
p51 <- ggscatter(plt_dt, x="cor1", y = "cor2", color = "protein", 
                label="protein", repel =TRUE)  + 
  xlab("Accuracy of initilized Z") + ylab("Accuracy of imputed Z")  + 
  theme(legend.position = "none") + xlim(0,1) + ylim(0,1) + 
  geom_abline(intercept = 0, slope = 1, linetype="dashed")
 p51
 
 
  
```

# show key protein markers change after bindsc alignment 
```{r, message = FALSE, warning=FALSE, results='hide',fig.height=8, fig.width=16}
marker_lst <- c("CD19","CD14","CD11c","CD28")
True <- X[marker_lst,]
init <- Z0[marker_lst,]
impu <- res$Z_est[marker_lst,]
p_lst <-c()
for(i in seq(1, length(marker_lst),1)){
  True[i,] <- True[i,]/max(True[i,])
  init[i,] <- init[i,]/max(init[i,])
  plt_dt <- data.frame("True"=True[i,], "Est"=init[i,])
  p61 <- ggplot(plt_dt, aes(x=True, y=Est) ) + geom_bin2d() + theme_classic() + 
  xlab("True protein level") + ylab("Initilized gene score matrix") + 
  ggtitle(marker_lst[i])+ 
  scale_fill_gradient(low="lightblue1",high="darkblue",trans="log10") 
  
  
  impu[i,] <- (impu[i,]-min(impu[i,]))/max(impu[i,])
  plt_dt <- data.frame("True"=True[i,], "Est"=impu[i,])
  p62 <- ggplot(plt_dt, aes(x=True, y=Est) ) + geom_bin2d() + theme_classic() + 
  xlab("True protein level") + ylab("Imputed gene score matrix") + 
  ggtitle(marker_lst[i])+
  scale_fill_gradient(low="lightblue1",high="darkblue",trans="log10") 
  p_lst[[i*2-1]] <- p61
  p_lst[[2*i]] <- p62
  
}

p6 <- ggarrange(p_lst[[1]],p_lst[[3]],p_lst[[5]],p_lst[[7]],
                p_lst[[2]],p_lst[[4]],p_lst[[6]],p_lst[[8]],ncol=4, nrow=2)
print(p6)


```
```{r save results }
save.image(file="PBMC.3k.RDS")
sessionInfo()
```