---
title: "Testing bindSC integration using simulated data"
author: "Jinzhuang Dou"
date: "July 7, 2020"
output: html_document
---

```{r setup, include=FALSE,cache=TRUE}
knitr::opts_chunk$set(echo = TRUE)
```

This guide will demonstrate the usage of bindSC to intergrate RNA and ATAC under the gene regulation modeling with  expression level regulated by both proximal and distal peaks. 
   
### Stage 1: Generate simulation  data
User can refer to `sim.R` for the details of simulaiton process (It will take 5 mins). 

```{r load data,echo=TRUE, warning=FALSE, message=FALSE}
library(bindSC)
library(ggpubr)
source("./sim.R")

prob_vec <- c(0.3,0.3,0.4)  # three populaitons 
w <- 0.5   # contribution of distal peaks on gene regulation 
nGene <- 500
nPeak <- 1500
nCell <- 2000

sim<- sim_XYZ(cell_perct= prob_vec, w=w, nGene=nGene, nPeak=nPeak, nCell=nCell)



dim(sim$X)
dim(sim$Y)
dim(sim$Z0)
```

Usage of BiCCA 

```{r usage, warning=TRUE, message=TRUE}

?BiCCA

```

In this example, dataset `X` is the gene expression matrix and `Y` is the peak matrix.  `Z0` is the simulated gene activity matrix.  Seurat, Liger and Harmony use (`X`, `Z0`) as input and BiCCA uses (`X`, `Z0`, `Y`) as input. 

bindSC also requires the modality specfic clustering ID `X.clst`,`Y.clst` to test the integraiton performance
 
The option `tolereance` is usually set from 0.01 to 0.05 to avoid unnecessary iteration when sample size is large.

```{r run BiCCA}
out <- BiCCA(X=sim$X, Z0=sim$Z0, Y=sim$Y, 
      K = 5,
      alpha = 0.1,
      lambda = 0.5,
      X.clst = sim$x.clst,
      Y.clst = sim$y.clst,  
      num.iteration = 100, 
      temp.path = "./out",
      tolerance = 0.01, 
      save = TRUE, 
      block.size =10000)

```

Show the iteration index `delta` 

```{r output check1, fig.height=4, fig.width=4}
plot(out$delta, xlab="step", ylab="delta")
```


There are three key ouputs.The vectors `u` and `v` are the low-dimension reprentation of two modalites which can be used for downstream joint clustering. The matrix `Z_est` is the updated gene score matrix after iterations. 

```{r output check}
summary(out)
```

### Stage 2: Comparsion between bindSC and the traditonal CCA

```{r method comparison, echo=TRUE, warning=FALSE, fig.height=5, fig.width=6}
cell_type <- c(sim$x.clst, sim$y.clst)
data_type <- c(rep("A", dim(out$u)[1]), rep("B",dim(out$r)[1]))

bindsc_umap <- umap(rbind(out$u, out$r))

plt_dt <- data.frame("UMAP1"=bindsc_umap$layout[,1],
                     "UMAP2"=bindsc_umap$layout[,2],
                     "cell_type"= as.factor(cell_type),
                     "data_type"=data_type)

p<-ggscatter(plt_dt, x = "UMAP1", y = "UMAP2",
     color = "cell_type", palette = c("darkseagreen4","lightpink","darkorchid1"), 
     repel = FALSE,size=1, alpha=1,legend.title = "", title="bindSC", font.title=16) + facet_grid(~data_type) 
p

```

Performance on traditional CCA method (results are from the first iteration). 

```{r method comparison cca, echo=TRUE, warning=FALSE, fig.height=5, fig.width=6}
out <- readRDS(file="./out/iteration1.rds")
cca_umap <- umap(rbind(out$u, out$r))

plt_dt <- data.frame("UMAP1"=cca_umap$layout[,1],
                     "UMAP2"=cca_umap$layout[,2],
                     "cell_type"= as.factor(cell_type),
                     "data_type"=data_type)
# remove some outliers from CCA
p<-ggscatter(plt_dt[(plt_dt$UMAP1>(-20) & plt_dt$UMAP2<20), ], x = "UMAP1", y = "UMAP2",
     color = "cell_type", palette = c("darkseagreen4","lightpink","darkorchid1"), 
     repel = FALSE,size=1, alpha=1,legend.title = "", title="CCA", font.title=16) + facet_grid(~data_type)
p

```
```{r summary}
sessionInfo()
```
