---
title: "Testing unmatched feature integration using simulated data"
author: "Jinzhuang Dou"
date: "July 7, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This guide will demonstrate the usage of bindSC to intergrate two datasets with unmatched features from simulation dataset. 
   
### Stage 1: Load data
**For convenience, we have prepared the pre-processed data which are ready to use.** 
User can refer to `sim.R`  for the details of running the pre-processing workflow (It will take 10 mins). 

```{r load data,echo=TRUE, warning=TRUE, message=TRUE}
library(bindSC)
#source("../../R/BiCCA.R")
#source("../../R/generics.R")


dim(sim$X)
dim(sim$Y)
dim(sim$Z0)
```

Usage of BiCCA 

```{r usage, warning=TRUE, message=TRUE}

?BiCCA

```
In this example, dataset `X` is the same with `Y`. The inital transiton matrix `Z0` is generated by permutating rows of `X` with 10% probablity the same with `X`.  Seurat and Liger use (`X`, `Z0`) as input and BiCCA uses (`X`, `Z0`, `Y`) as input. 
 
The option `tolereance` is usually set from 0.01 to 0.05 to avoid unnecessary iteration when sample size is large. Here I set it to 0.001 for small sample size. 

```{r run BiCCA}

out <- BiCCA(X=sim$X,
             Z0=sim$Z0, 
             Y=sim$Y, 
             alpha = 0,
             lambda = 0.5,
             K = 5, 
             X.clst = sim$X_meta$Group,
             Y.clst = sim$Y_meta$Group,
             num.iteration = 100, 
             temp.path = "./tp",
             tolerance = 0.01, 
             save = TRUE, 
             block.size = 5000)

```

```{r output check}
summary(out)
```
Show the iteration index `delta` 
```{r output check1, fig.height=4, fig.width=4}
plot(log10(out$delta), xlab="step", ylab="-log10(delta)")
```

### Stage 2: Comparsion among bindSC, Seurat, and Liger methods 

Three metrics are used to measure method performance since we have cell correspondence as the gold standard.
**For convenience, we have prepared the results for Seurat and Liger which are ready to use.** 

- `Silhoutte coefficient` : High value means cell-type architectures is well preserved in latent space
- `Alignment score` : High value means uniformity of mixing for two datasets in latent space
- `Anchor accuracy` : High value means cell correspondence can be found in cell's neighbor given fixed neighbor size


```{r method comparison, echo=TRUE, warning=FALSE, }
source("../A549/eval_plot.r")
cell_type <- c(sim$X_meta$Group, sim$Y_meta$Group)
data_type <- c(rep("A", dim(out$u)[1]), rep("B",dim(out$r)[1]))
myumap <- umap(rbind(out$u, out$r))
result <- umap_plot(myumap, cell_type, data_type)
eval <-  method_compare(result$plt_dt, "Sim")

```
```{r plot1, echo=TRUE, fig.height=5, fig.width=8}
eval$coembedding

```

For anchor accuracy, the gray line denotes results from random guess. 

```{r plot2, echo=TRUE, fig.height=4, fig.width=8}
eval$eval
```


